D3064R0: Making OOTA Execution on Real Implementations be Vacuous

### 2024-01-06 ootavacuous.tex
Cross-linking, formatting, and wordsmithing
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{document}
\title{D3064R0: OOTA Execution on Real Implementations is Provably Vacuous}

@@

The current authors deviate slightly from this tradition by instead
defining an OOTA cycle as a cycle in sdep $\cup$ rfe (or
\co{acyclic(sdep ; rfe)} in \co{herd7} notation), where rfe is the
set of store-to-load links, but excluding intra-thread rfi links, which
are instead considered to be part of sdep.
Although this does shunt even more complexity onto the term
``semantic dependency'', it also enables us to bring to bear
laws of physics and properties of hardware onto correct real-world
C++ implementations.% \footnote{
%	Thanks to Alan Stern for providing a litmus test that demonstrated
%	that we had unwittingly (but productively!) shifted our definition
%	to sdep $\cup$ rfe.}

@@

@@

@@

This situation underscores the point that the existance of a semantic
dependency can be a property of a particular execution rather than a
universal property of the code itself.

@@

~\\
{
	\scriptsize
	\texttt{
	gcc -Wp,-MMD,kernel/rcu/.tree.o.d -nostdinc
	-I./arch/x86/include -I./arch/x86/include/generated  -I./include
	-I./arch/x86/include/uapi -I./arch/x86/include/generated/uapi
	-I./include/uapi -I./include/generated/uapi
	-include ./include/linux/compiler-version.h
	-include ./include/linux/kconfig.h -include
	./include/linux/compiler\_types.h -D\_\_KERNEL\_\_
	-fmacro-prefix-map=./= -Werror -std=gnu11 -fshort-wchar
	-funsigned-char -fno-common -fno-PIE -fno-strict-aliasing -mno-sse
	-mno-mmx -mno-sse2 -mno-3dnow -mno-avx -fcf-protection=branch
	-fno-jump-tables -m64 -falign-jumps=1 -falign-loops=1
	-mno-80387 -mno-fp-ret-in-387 -mpreferred-stack-boundary=3
	-mskip-rax-setup -mtune=generic -mno-red-zone -mcmodel=kernel
	-Wno-sign-compare -fno-asynchronous-unwind-tables
	-mindirect-branch=thunk-extern -mindirect-branch-register
	-mindirect-branch-cs-prefix -mfunction-return=thunk-extern
	-fno-jump-tables -fpatchable-function-entry=16,16
	-fno-delete-null-pointer-checks -O2 -fno-allow-store-data-races
	-fstack-protector-strong -fomit-frame-pointer
	-fno-stack-clash-protection -falign-functions=16
	-fno-strict-overflow -fno-stack-check -fconserve-stack
	-Wall -Wundef -Werror=implicit-function-declaration
	-Werror=implicit-int -Werror=return-type -Werror=strict-prototypes
	-Wno-format-security -Wno-trigraphs -Wno-frame-address
	-Wno-address-of-packed-member -Wframe-larger-than=2048 -Wno-main
	-Wno-unused-but-set-variable -Wno-unused-const-variable -Wvla
	-Wno-pointer-sign -Wcast-function-type -Wno-array-bounds
	-Wno-alloc-size-larger-than -Wimplicit-fallthrough=5
	-Werror=date-time -Werror=incompatible-pointer-types
	-Werror=designated-init -Wenum-conversion
	-Wno-unused-but-set-variable -Wno-unused-const-variable
	-Wno-restrict -Wno-packed-not-aligned -Wno-format-overflow
	-Wno-format-truncation -Wno-stringop-overflow
	-Wno-stringop-truncation -Wno-missing-field-initializers
	-Wno-type-limits -Wno-shift-negative-value
	-Wno-maybe-uninitialized -Wno-sign-compare
	-DKBUILD\_MODFILE='"kernel/rcu/tree"' -DKBUILD\_BASENAME='"tree"'
	-DKBUILD\_MODNAME='"tree"' -D\_\_KBUILD\_MODNAME=kmod\_tree -c
	-o kernel/rcu/tree.o kernel/rcu/tree.c
	}
}

@@

% @@@ TODO:
% Construct interleaved OOTA examples (see sheets of paper).
% Check cmbc on nondeterministic i++.
% Use LKMM as check for dependencies?  Check invented-load litmus tests.

>>>>

\begin{document}
\title{D3064R0: Making OOTA Execution on Real Implementations be Vacuous}

@@

The current authors deviate slightly from this tradition by instead
defining an OOTA cycle as a cycle in sdep $\cup$ rfe (or
\co{acyclic(sdep ; rfe)} in \co{herd7} notation), where rfe is the
set of store-to-load links, but excluding intra-thread rfi links, which
are instead considered to be part of sdep.
Although this does shunt even more complexity onto the term
``semantic dependency'', it also enables us to bring to bear
laws of physics and properties of hardware onto correct real-world
C++ implementations.
% \footnote{
%	Thanks to Alan Stern for providing a litmus test that demonstrated
%	that we had unwittingly (but productively!) shifted our definition
%	to sdep $\cup$ rfe.}
Please see
Appendix~\ref{app:Informal Definition of Semantic Dependency}
for more a more detailed (albeit quite informal) definition.

@@

It is also important to note that implementations and users have a
surprising degree of influence over language
semantics~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2}.
Please see
Appendix~\ref{app:User Influence Over Language Semantics}
for examples.

@@

For more discussion along these lines, please see
Section~\ref{sec:But What About Non-Volatile Atomics?}.

@@

This situation underscores the point that the existance of a semantic
dependency can be a property of a particular execution rather than a
universal property of the code itself.
The interested reader is encouraged to create more complex examples,
for example, demonstrating semantic dependencies from multiple loads or
exploring indeterminacy.

@@

~\\
{
	\scriptsize
	\texttt{
	gcc -Wp,-MMD,kernel/rcu/.tree.o.d -nostdinc
	-I./arch/x86/include \\
	-I./arch/x86/include/generated  -I./include
	-I./arch/x86/include/uapi \\
	-I./arch/x86/include/generated/uapi
	-I./include/uapi -I./include/generated/uapi \\
	-include ./include/linux/compiler-version.h
	-include ./include/linux/kconfig.h \\
	-include ./include/linux/compiler\_types.h -D\_\_KERNEL\_\_
	-fmacro-prefix-map=./= \\
	-Werror -std=gnu11 -fshort-wchar
	-funsigned-char -fno-common -fno-PIE \\
	-fno-strict-aliasing -mno-sse
	-mno-mmx -mno-sse2 -mno-3dnow -mno-avx \\
	-fcf-protection=branch
	-fno-jump-tables -m64 -falign-jumps=1 -falign-loops=1 \\
	-mno-80387 -mno-fp-ret-in-387 -mpreferred-stack-boundary=3
	-mskip-rax-setup \\
	-mtune=generic -mno-red-zone -mcmodel=kernel
	-Wno-sign-compare \\
	-fno-asynchronous-unwind-tables
	-mindirect-branch=thunk-extern \\
	-mindirect-branch-register
	-mindirect-branch-cs-prefix -mfunction-return=thunk-extern
	-fno-jump-tables -fpatchable-function-entry=16,16
	-fno-delete-null-pointer-checks -O2 \\
	-fno-allow-store-data-races
	-fstack-protector-strong -fomit-frame-pointer \\
	-fno-stack-clash-protection -falign-functions=16
	-fno-strict-overflow -fno-stack-check \\
	-fconserve-stack
	-Wall -Wundef -Werror=implicit-function-declaration \\
	-Werror=implicit-int -Werror=return-type -Werror=strict-prototypes \\
	-Wno-format-security -Wno-trigraphs -Wno-frame-address
	-Wno-address-of-packed-member \\
	-Wframe-larger-than=2048 -Wno-main
	-Wno-unused-but-set-variable \\
	-Wno-unused-const-variable -Wvla
	-Wno-pointer-sign -Wcast-function-type \\
	-Wno-array-bounds
	-Wno-alloc-size-larger-than -Wimplicit-fallthrough=5 \\
	-Werror=date-time -Werror=incompatible-pointer-types
	-Werror=designated-init \\
	-Wenum-conversion
	-Wno-unused-but-set-variable -Wno-unused-const-variable \\
	-Wno-restrict -Wno-packed-not-aligned -Wno-format-overflow
	-Wno-format-truncation \\
	-Wno-stringop-overflow
	-Wno-stringop-truncation -Wno-missing-field-initializers \\
	-Wno-type-limits -Wno-shift-negative-value
	-Wno-maybe-uninitialized \\
	-Wno-sign-compare
	-DKBUILD\_MODFILE='"kernel/rcu/tree"'
	-DKBUILD\_BASENAME='"tree"' \\
	-DKBUILD\_MODNAME='"tree"' -D\_\_KBUILD\_MODNAME=kmod\_tree \\
	-c -o kernel/rcu/tree.o kernel/rcu/tree.c
	}
}

@@

% @@@ TODO:
% Construct interleaved OOTA examples (see sheets of paper).
% Check cmbc on nondeterministic i++.
% Use LKMM as check for dependencies?
